Microlensing as a possible probe of event-horizon structure in quasars

### 2016-10-10 paper1/ms.tex

The gravitational lens equation
\begin{equation}
\vec\beta = \vec\theta - \vec\alpha(\vec\theta)
\label{eqn:lens}
\end{equation}
relates the apparent sky position $\vec\theta$ of a light source to
its true but unobservable sky position $\vec\beta$ through the bending
angle $\vec\alpha$.  The latter is an integral 
\begin{equation}
\vec\alpha = (1+z_{_L})\frac{D_S-D_L}{D_SD_L} \frac{4G}{c^2}
\int \Sigma(\vec\theta')
\frac{\vec\theta-\vec\theta'}{|\vec\theta-\vec\theta'|^2}\,
d^2\vec\theta'
\label{eqn:alpha}
\end{equation}
depending on the projected density $\Sigma(\vec\theta)$ (kg/steradian)
of the lens, the lens redshift $z_{_L}$ and the comoving distances
$D_L$ and $D_S$ to the lens and source.  The derivative of the
apparent position with respect to the source position
\begin{equation}
M(\vec\theta) =
\left(\frac{\partial\beta}{\partial\theta}\right)^{-1}
\label{eqn:magnif-matrix}
\end{equation}
is known as the magnification matrix, and its determinant
\begin{equation}
\mu(\vec\theta) = \det|M(\vec\theta)|
\end{equation}
is the brightness amplification of an image of a point source.  In
other words, the source will brighten or dim according to whether
$\mu(\vec\theta)$ is more or less than unity.  If there are multiple
mages at distinct but not observationally resolved $\vec\theta_i$ from
the same $\vec\beta$, a total brightness amplification
of\begin{equation} \mu_{\rm total} = \sum_{i} \mu(\vec\theta_i)
\end{equation}
applies. It is possible for the magnification to become formally
infinite, as a result of an eigenvalues of the magnification matrix
(\ref{eqn:magnif-matrix}) becoming infinite.  This generically happens
on curves on the $\vec\theta$ plane, known as {\em critical curves}.
Mapping a critical curve to the source plane, through the lens
equation (\ref{eqn:lens}), to the source plane $\vec\beta$ gives the
so-called caustics.  Caustics can appear in the optical system, not just
gravitational lensing.  For a point source, caustics are singularities
of the magnification; for finite-size sources caustics correspond to
high and sharply changing magnification.

@@

\begin{equation}
 S_{2D}^{Dp}(p-p_{sn}, q-q_{sn}) = \frac{S_0^{Dp}}{\pi R_p^2} \Theta \left( R_p^2 - \left( p-p_{sp} \right)^2 - \left( q-q_{sp} \right)^2 \right)
\end{equation}
\\
and
\begin{equation}
 S_{2D}^{Dn}(p-p_{sn}, q-q_{sn}) = \frac{S_0^{Dn}}{\pi R_n^2} \Theta \left( R_n^2 - \left( p-p_{sn} \right)^2 - \left( q-q_{sn} \right)^2 \right).
\end{equation}
\\
The following notations were used: $R_p, (p_{sp}, q_{sp}), R_n, (p_{sn},q_{sn})$ are the radii and coordinate of the center for the larger positive disk and smaller negative disk respectively.  $S_0^{Dp},S_0^{Dn}$ represent the total flux of radiation received from the large and small disk. From this point forward we will not use the total flux from each source. Instead we will 
use the difference which in this case is th total flux from the crescent-shaped source $S_0^C$. \\
Equation \ref{eqn:s2d} can be written as:\\

@@

\begin{equation}
        p_3 -p_2 = 2 R_n.
\end{equation}
All four points mark the positions where derivative $\frac{dS_{1D}}{dp}$ is discontinuous. The points can be used to define three regions: $p_1 - p_2$ where $S_1D$ is convex, $p_2 - p_3$ where $S_1D$ is concave, and $p_3 - p_4$ where $S_1D$ is convex again. \\

@@

Once the map is created, the lightcurve can be obtained by specifying
the transit path of the source across the map.  The source is moved
along the transit line, one pixel at a time, and at each step the
brightness distribution of the source is convolved
(equation~\ref{eqn:ft2d}) with the magnification map to give the
brightness.  In real life, not only are both lens and source moving
but the lens lens configuration, and with it the magnification
pattern, is also changing with time.  While the first subtlety is
taken care of by a coordinate transformation in this analysis, for the
second one the lens configuration is assumed to be constant in time.

>>>>

The gravitational lens equation
\begin{equation}
\vec\beta = \vec\theta - \vec\alpha(\vec\theta)
\label{eqn:lens}
\end{equation}
relates the apparent sky position $\vec\theta$ of a light source to
its true but unobservable sky position $\vec\beta$ through the bending
angle $\vec\alpha$.  The latter is an integral 
\begin{equation}
\vec\alpha = (1+z_{_L})\frac{D_S-D_L}{D_SD_L} \frac{4G}{c^2}
\int \Sigma(\vec\theta')
\frac{\vec\theta-\vec\theta'}{|\vec\theta-\vec\theta'|^2}\,
d^2\vec\theta'
\label{eqn:alpha}
\end{equation}
depending on the projected density $\Sigma(\vec\theta)$ (kg/steradian)
of the lens, the lens redshift $z_{_L}$ and the comoving distances
$D_L$ and $D_S$ to the lens and source.  The derivative of the
apparent position with respect to the source position
\begin{equation}
M(\vec\theta) =
\left(\frac{\partial\beta}{\partial\theta}\right)^{-1}
\label{eqn:magnif-matrix}
\end{equation}
is known as the magnification matrix, and its determinant
\begin{equation}
\mu(\vec\theta) = \det|M(\vec\theta)|
\end{equation}
is the brightness amplification of an image of a point source.  In
other words, the source will brighten or dim according to whether
$\mu(\vec\theta)$ is more or less than unity.  If there are multiple
images at distinct but not observationally resolved $\vec\theta_i$ from
the same $\vec\beta$, a total brightness amplification
of

\begin{equation} \mu_{\rm total} = \sum_{i} \mu(\vec\theta_i)
\end{equation}
applies. It is possible for the magnification to become formally
infinite, as a result of an eigenvalues of the magnification matrix
(\ref{eqn:magnif-matrix}) becoming infinite.  This generically happens
on curves on the $\vec\theta$ plane, known as {\em critical curves}.
Mapping a critical curve to the source plane, through the lens
equation (\ref{eqn:lens}), to the source plane $\vec\beta$ gives the
so-called caustics.  Caustics can appear in the optical system, not just
gravitational lensing.  For a point source, caustics are singularities
of the magnification; for finite-size sources caustics correspond to
high and sharply changing magnification.

@@

\begin{equation}
 S_{2D}^{Dp}(p-p_{sn}, q-q_{sn}) = \frac{S_0^{Dp}}{\pi R_p^2} \Theta \left( R_p^2 - \left( p-p_{sp} \right)^2 - \left( q-q_{sp} \right)^2 \right)
\end{equation}
\\
and
\begin{equation}
 S_{2D}^{Dn}(p-p_{sn}, q-q_{sn}) = \frac{S_0^{Dn}}{\pi R_n^2} \Theta \left( R_n^2 - \left( p-p_{sn} \right)^2 - \left( q-q_{sn} \right)^2 \right).
\end{equation}
\\
The following notations were used: $R_p, (p_{sp}, q_{sp}), R_n, (p_{sn},q_{sn})$ are the radii and coordinate of the center for the larger positive disk and smaller negative disk respectively.  $S_0^{Dp},S_0^{Dn}$ represent the total flux of radiation received from the large and small disk. From this point forward we will not use the total flux from each source. Instead we will 
use the difference which in this case is the total flux from the crescent-shaped source $S_0^C$. \\
Equation \ref{eqn:s2d} can be written as:\\

@@

\begin{equation}
        p_3 -p_2 = 2 R_n.
\end{equation}
All four points mark the positions where derivative $\frac{dS_{1D}}{dp}$ is discontinuous. The points can be used to define three regions: $p_1 - p_2$ where $S_{1D}$ is convex, $p_2 - p_3$ where $S_{1D}$ is concave, and $p_3 - p_4$ where $S_{1D}$ is convex again. \\

@@

Once the map is created, the lightcurve can be obtained by specifying
the transit path of the source across the map.  The source is moved
along the transit line, one pixel at a time, and at each step the
brightness distribution of the source is convolved
(equation~\ref{eqn:ft2d}) with the magnification map to give the
brightness.  In real life, not only are both lens and source moving
but the lens configuration, and with it the magnification
pattern, is also changing with time.  While the first subtlety is
taken care of by a coordinate transformation in this analysis, for the
second one the lens configuration is assumed to be constant in time.
